---
title: "Export to DATSU"
author: "Cedric"
format: 
  html :
    self-contained: true
editor: visual
params:
  samplingscheme: 
  - "GiTCG"
  - "VacG"
  - "BresGY"  
  - "BreS"  
  species: "eel"
---

```{r }
#| label: qmdhelper 
#| echo: FALSE
#| eval: FALSE

# this is only necessary in eclipse, Rstudio or vscodium should down
params <- list(species = "eel",
samplingscheme=c("GiTCG", "VacG", "BresGY", "BreS")

)
```

```{r init}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'
#if (!grepl("montepomi", getwd())) {
if(Sys.info()[["user"]] == 'joliviero'){
  setwd("D:/workspace/DIASPARA_WP3_migdb/R")
  datawd <- "D:/DIASPARA/wgbast"
} else if (Sys.info()[["user"]] == 'cedric.briand'){
  setwd("C:/workspace/DIASPARA_WP3_migdb/R")
  datawd <- "C:/Users/cedric.briand/OneDrive - EPTB Eaux&Vilaine/Projets/DIASPARA/wgbast"
}
source("utilities/load_library.R")
load_library("tidyverse")
load_library("knitr")
load_library("kableExtra")
load_library("RPostgres")
load_library("yaml")
load_library("uuid")
load_library("glue")
cred <- read_yaml("../credentials.yml")
# check if internal

test_internal <- function(host) {
  status <- system2("ping", args = c(" 10.1.128.16"), stdout = FALSE, stderr = FALSE)
  return(status == 0)
}


if (test_internal()){
con_diaspara <- dbConnect(Postgres(), 
                          dbname = cred$dbnamediaspara,
                          host = cred$hostinternal,
                          port = cred$port,
                          user = cred$userdiaspara,
                          password = cred$passworddiaspara)
} else {
con_diaspara <- dbConnect(Postgres(), 
                          dbname = cred$dbnamediaspara,
                          host = cred$host,
                          port = cred$port,
                          user = cred$userdiaspara,
                          password = cred$passworddiaspara)
}
folder <- tempdir()
schema <- paste0("dat",params$species)
refschema <- paste0("ref",params$species)
```

TODO check with Maria : DatarightsHolder doi Station code

```{r}
#| label: build_dataset_FI
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'

# File info
# re

query <- 
glue_sql('SELECT \'FI\' AS "RecordType",ser_cou_code AS "Country", 
ver_year AS "Year",
ser_spe_code AS "Species",
ser_wkg_code AS "WorkingGroup",
ver_version AS "Version",
1001 AS "DatarightsHolder" 
FROM {`schema`}.t_series_ser
JOIN {`refschema`}.tr_version_ver on ver_code = ser_ver_code
WHERE ser_code IN ({ss*}) LIMIT 1',
  ss = params$samplingscheme, .con = con_diaspara)

FI <- dbGetQuery(con_diaspara, query)

write.csv(FI,
          file.path(folder, "FI.csv"),
                    quote =FALSE,           na = "",           row.names = FALSE)
```

FI : `r nrow(FI)`

```{r}
#| label: build_dataset_sampling_scheme_SS
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'
query <- 
glue_sql('SELECT 
\'SS\' AS "RecordType",
ser_code AS "SamplingSchemeID",
10777 AS "StationCode", 
ser_fiw_code AS "Fishway",
ser_mon_code AS "MonitoringDevice",
ser_spe_code AS "Species",
ser_lfs_code AS "LifeStage",
ser_are_code AS "Area",
gea_icesvalue AS "Gear",
case WHEN ser_stocking THEN \'Y\'
     WHEN NOT ser_stocking THEN \'N\' 
     ELSE NULL
     END AS "RestockingFlag",
\'https://doi.org/10.17895/ices.pub.30488120\' AS DOI,
ser_gea_code,
gea_description
FROM {`schema`}.t_series_ser
JOIN ref.tr_gear_gea on gea_code = ser_gea_code
WHERE ser_code IN ({ss*})',
ss = params$samplingscheme, .con = con_diaspara)


SS <- dbGetQuery(con_diaspara, query)

# TODO REMOVE LATER: Fix because some gears don't yet have a correspondence

missing_gears <- paste(SS$SamplingSchemeID[is.na(SS$Gear)],
                       SS$ser_gea_code[is.na(SS$Gear)],
                       SS$gea_description[is.na(SS$Gear)]) |>  purrr::discard(is.na) 
SS <- SS |> select(-ser_gea_code, -gea_description)


write.csv(SS,
          file.path(folder, "SS.csv"),
                    quote =FALSE,        
                    na = "",
                    row.names = FALSE)
```

Missing gears : `r missing_gears` TODO check the definition of proportion G in GY Y in YS becomes lifestageProportion.

TODO comments attached to all group metrics ? Or comment attached to each metric.

SS : `r nrow(SS)`

```{r}
#| label: build_dataset_group_metrics_GM
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'



# GM GROUP METRICS 
# note sex is in subsamples not samples

query <- 
glue_sql('
SELECT 
\'GM\' AS "RecordType",
gr_id AS "GroupID",
ser_code AS "SamplingSchemeID",
gr_year AS "Year",
NULL AS "AgeClass",
gr_lfs_code AS "LifeStage",
gr_number AS "NumberofIndividuals",
gr_comment AS "Comments",            
-- gr_sex_code AS "Sex",
grt_tra_code,
grt_value,
grt_trm_code,
grt_trv_code
FROM {`schema`}.t_series_ser 
JOIN {`schema`}.t_group_gr ON gr_ser_id = ser_id 
JOIN {`schema`}.t_grouptrait_grt ON grt_gr_id = gr_id 
AND gr_gr_id IS NULL',
ss = params$samplingscheme, .con = con_diaspara)
GM0 <- dbGetQuery(con_diaspara, query)

# GM grt_tra_code, grt_value, grt_trv_code, grt_trm_code

GMnumeric <-
  GM0 |> 
  filter(is.na(grt_trv_code) & !is.na(grt_value))  |>
  select(-grt_trv_code) 

GMqualitative <-   GM0 |> 
  filter(!is.na(grt_trv_code) & is.na(grt_value))  |>
  select(-grt_value) |>
    pivot_wider(
  names_from = all_of(c("grt_tra_code")), 
  values_from ="grt_trv_code") 

# no qualitative values for Group metrics

GM_without_method <- GMnumeric |> filter(is.na(grt_trm_code)) |>
  select(-grt_trm_code) |>
  pivot_wider(
  names_from = all_of(c("grt_tra_code")), 
  values_from = all_of(c("grt_value")))
  
GM_with_method <- GMnumeric |> filter(!is.na(grt_trm_code)) |>
  pivot_wider(
  names_from = all_of(c("grt_tra_code")), 
  values_from = all_of(c("grt_value", "grt_trm_code"))) |>
  rename("Muscle_lipid" = "grt_value_Muscle_lipid",
         "Anguillicola_proportion" = "grt_value_Anguillicola_proportion",
         "Anguillicola_intensity" = "grt_value_Anguillicola_intensity",
         "Female_proportion"="grt_value_Female_proportion",
         "MuscleLipidMethod"="grt_trm_code_Muscle_lipid",
         "SexDeterminationMethod"="grt_trm_code_Female_proportion",
         "AnguillicolaMethod"="grt_trm_code_Anguillicola_intensity") |>
  select (-grt_trm_code_Anguillicola_proportion) 
# last one grt_trm_code_Anguillicola_proportion is twice there, the same for
# proportion and intensity

# GM_structure used to get all columns for the format
GM_structure <- data.frame(
  RecordType = character(),
  GroupID = character(),
  SamplingSchemeID = character(),
  Year = integer(),
  AgeClass = character(),
  LifeStage = character(),
  LifeStageProportion = numeric(),
  NumberofIndividuals = integer(),
  Age = numeric(),
  Length = numeric(),
  Weight = numeric(),
  SexDeterminationMethod = character(),
  FemaleProportion = numeric(),
  DifferentiatedProportion = numeric(),
  AnguillicolaProportion = numeric(),
  AnguillicolaIntensity = numeric(),
  AnguillicolaMethod = character(),
  MuscleLipidPercent = numeric(),
  MuscleLipidMethod = character(),
  PCB = numeric(),
  TEQ = numeric(),
  EVEXProportion = numeric(),
  HVAProportion = numeric(),
  Pb = numeric(),
  Cd = numeric(),
  Hg = numeric(),
  Comments = character(),
  stringsAsFactors = FALSE
)

# need to merge proportions
# No data currently in group metrics for PCB or or TEQ

GM1 <- bind_rows(
  GM_without_method |> mutate(GroupID = as.character(GroupID)),
  GM_with_method |> mutate(GroupID = as.character(GroupID))
) |> mutate(LifeStageProportion = 
              ifelse(is.na(S_in_ys_proportion), G_in_gy_proportion, S_in_ys_proportion)) |>
  rename("MeanLength" = "Lengthmm",
         "MeanWeight" = "Weightg",
         "MeanAge" = "Ageyear",
         "FemaleProportion"="Female_proportion",
         "AnguillicolaProportion" = "Anguillicola_proportion",
         "DifferentiatedProportion" = "Differentiated_proportion",
         "AnguillicolaIntensity"="Anguillicola_intensity",
         "MuscleLipidPercentage"= "Muscle_lipid",
         "EVEXProportion" = "Evex_proportion",
         "HVAProportion"="Hva_proportion"
        )  


 GM <-  bind_rows(GM_structure , GM1) |>
    select(RecordType,GroupID,SamplingSchemeID,Year,AgeClass,LifeStage, 
          LifeStageProportion, NumberofIndividuals, MeanAge,  MeanWeight,  SexDeterminationMethod,  FemaleProportion,  DifferentiatedProportion,  AnguillicolaProportion,  AnguillicolaIntensity, AnguillicolaMethod,  MuscleLipidPercentage, MuscleLipidMethod,
          PCB, TEQ, EVEXProportion, HVAProportion,Pb,Cd,Hg, Comments)
 
 # " PCB" = "Sum_6_pcb",


GM <- GM |> 
  filter(SamplingSchemeID %in% params$samplingscheme) 

write.csv(GM,
          file.path(folder, "GM.csv"),
                    quote =FALSE,        
                    na = "",
                    row.names = FALSE)

```

CHECK check subgroup can be based on stage or sex, there should be a method to distinguish between stages. There is no comment with GS... But should there be one ? Probably if you follow the same format it would not add much ?

GM : `r nrow(GM)`

```{r}
#| label: build_dataset_group_subsample_GS
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'

# GS note sex is in subsamples not samples
query <- 
glue_sql('
SELECT 
\'GM\' AS "RecordType",
gr_id AS "GroupID",
ser_code AS "SamplingSchemeID",
gr_year AS "Year",
NULL AS "AgeClass",
gr_lfs_code AS "LifeStage",
gr_number AS "NumberofIndividuals",
gr_comment AS "Comments",            
gr_sex_code AS "Sex",
grt_tra_code,
grt_value,
grt_trm_code,
grt_trv_code
FROM {`schema`}.t_series_ser 
JOIN {`schema`}.t_group_gr ON gr_ser_id = ser_id 
JOIN {`schema`}.t_grouptrait_grt ON grt_gr_id = gr_id 
AND gr_gr_id IS NOT NULL', .con = con_diaspara)
GS0 <- dbGetQuery(con_diaspara, query)

# GM grt_tra_code, grt_value, grt_trv_code, grt_trm_code

GSnumeric <-
  GS0 |> 
  filter(is.na(grt_trv_code) & !is.na(grt_value))  |>
  select(-grt_trv_code) 

GSqualitative <-   GS0 |> 
  filter(!is.na(grt_trv_code) & is.na(grt_value))  |>
  select(-grt_value) |>
    pivot_wider(
  names_from = all_of(c("grt_tra_code")), 
  values_from ="grt_trv_code") 

# no qualitative values for Group metrics

GS_without_method <- GSnumeric |> filter(is.na(grt_trm_code)) |>
  select(-grt_trm_code) |>
  pivot_wider(
  names_from = all_of(c("grt_tra_code")), 
  values_from = all_of(c("grt_value")))

# method is always about sex there   
GS_with_method <- GSnumeric |> filter(!is.na(grt_trm_code)) |>
  pivot_wider(
  names_from = all_of(c("grt_tra_code")), 
  values_from = all_of(c("grt_value"))) |>
  rename("SexDeterminationMethod" = "grt_trm_code") 

# last one grt_trm_code_Anguillicola_proportion is twice there, the same for
# proportion and intensity
# 
GS_structure <- data.frame(
  RecordType = character(),            # char(2)
  SubSampleID = character(),           # char(10)
  GroupID = character(),               # char(20)
  Sex = character(),                   # char(2), code list SEXCO
  LifeStage = character(),             # char(10), code list STAGE
  Age = integer(),                     # int(2)
  Length = numeric(),                  # float(10)
  Weight = numeric(),                  # float(10)
  AnguillicolaProportion = numeric(),  # float(5)
  AnguillicolaIntensity = numeric(),   # float(5)
  AnguillicolaMethod = character(),    # char(20), code list METOA
  MuscleLipidPercent = numeric(),      # float(10)
  MuscleLipidMethod = character(),     # char(20), code list METOA
  Comments = character(),              # char(200)
  stringsAsFactors = FALSE
)
 
GS <- bind_rows(
  GS_without_method,
  GS_with_method) |>
  rename(Length = Lengthmm,
         Weight = Weightg,
         Age = Ageyear)  |>
    select(RecordType,GroupID,SamplingSchemeID,Year,AgeClass,LifeStage, 
           NumberofIndividuals, Age,  Weight, Length,  SexDeterminationMethod)

GS <- bind_rows(GS_structure , GS|> mutate(GroupID = as.character(GroupID)))

GS <- GS |> 
  filter(SamplingSchemeID %in% params$samplingscheme) 

write.csv(GS,
          file.path(folder, "GS.csv"),
                    quote =FALSE,        
                    na = "",
                    row.names = FALSE)

```

Fish_ID =\> this is the identifier coming from country level ? Note : in our dataset, the method of sex is never used. Note : in our dataset there is no sex in individual metrics ... Note : anguillicola_presence not reported Note : I don't have TEQ Note : I don't have EVEX or HVA (check...)

GS : `r nrow(GS)`

```{r}
#| label: build_dataset_Individual_metrics_IM
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'

# GS note sex is in subsamples not samples
query <- 
glue_sql('
SELECT 
\'IM\' AS "RecordType",
fi_id,
CASE WHEN fi_date IS NULL THEN fi_year 
     ELSE fi_year END AS "Year",
ser_code AS "SamplingSchemeID",
fi_idsource AS "FishID",
fi_lfs_code AS "LifeStage",
CASE WHEN fi_date IS NULL THEN NULL 
ELSE extract(\'MONTH\' FROM fi_date)
END AS "Month",
CASE WHEN fi_date IS NULL THEN NULL 
ELSE extract(\'DAY\' FROM fi_date)
END AS "Day",
fi_x_4326 AS "DecimalLongitude",
fi_y_4326 AS "DecimalLatitude",
int_tra_code,
int_value,
int_trv_code,
int_trm_code,
fi_comment AS "Comments"
FROM {`schema`}.t_series_ser 
JOIN {`schema`}.t_fish_fi ON fi_ser_id = ser_id 
JOIN {`schema`}.t_indivtrait_int ON int_fi_id=fi_id',
.con = con_diaspara)
IM0 <- dbGetQuery(con_diaspara, query) # 1796871

# GM grt_tra_code, grt_value, grt_trv_code, grt_trm_code

IMnumeric <-
  IM0 |> 
  filter(is.na(int_trv_code) & !is.na(int_value))  |>
  select(-int_trv_code) 

IMqualitative <-   IM0 |> 
  filter(!is.na(int_trv_code) & is.na(int_value))  |>
  select(-int_value) |>
    pivot_wider(
  names_from = all_of(c("int_tra_code")), 
  values_from ="int_trv_code") 

# no qualitative values for Group metrics

IM_without_method <- IMnumeric |> filter(is.na(int_trm_code)) |>
  select(-int_trm_code) |>
  pivot_wider(
  names_from = all_of(c("int_tra_code")), 
  values_from = all_of(c("int_value")))

# method is always about sex there   
IM_with_method <- IMnumeric |> filter(!is.na(int_trm_code)) |>
  pivot_wider(
  names_from = all_of(c("int_tra_code")), 
  values_from = all_of(c("int_value", "int_trm_code"))) |>
  rename("MuscleLipidPercent"= "int_value_Muscle_lipid",
         "Anguillicola_intensity" = "int_value_Anguillicola_intensity",
         "MuscleLipidMethod"="int_trm_code_Muscle_lipid",
         "AnguillicolaMethod"="int_trm_code_Anguillicola_intensity") 

# last one int_trm_code_Anguillicola_proportion is twice there, the same for
# proportion and intensity


IMstructure <- data.frame(
  RecordType = character(),
  Year = integer(),
  SamplingSchemeID = character(),
  FishID = character(),
  LifeStage = character(),
  Month = integer(),
  Day = integer(),
  DecimalLongitude = numeric(),
  DecimalLatitude = numeric(),
  Sex = character(),
  SexDeterminationMethod = character(),
  Differentiated = character(),
  Age = integer(),
  SeaAge = integer(),
  JuvenileAge = integer(),
  AgingMethod = character(),
  AgingMatrix = character(),
  AgingPrepMethod = character(),
  Length = numeric(),
  LengthMethod = character(),
  Weight = numeric(),
  MeanEyeDiam = numeric(),
  PectoralFinLength = numeric(),
  AnguillicolaMethod = character(),
  AnguillicolaPresence = character(),
  AnguillicolaIntensity = numeric(),
  MuscleLipidPercent = numeric(),
  MuscleLipidMethod = character(),
  PCB = numeric(),
  TEQ = numeric(),
  EVEXpresence = character(),
  HVAPresence = character(),
  PigmentationStage = character(),
  Pb = numeric(),
  Cd = numeric(),
  Hg = numeric(),
  Comments = character(),
  stringsAsFactors = FALSE
)

# clipr::write_clip(colnames(IM))



 
IM <- bind_rows(
  IM_without_method,
  IM_with_method)  |>
mutate(FishID = ifelse(is.na(FishID), as.character(fi_id),FishID)) |>
  rename(
PectoralFinLength =Pectoral_lengthmm,
MeanEyeDiam=Eye_diam_meanmm,
Weight=Weightg,
Length=Lengthmm,
AnguillicolaIntensity = Anguillicola_intensity,
Age = Ageyear,
PCB = Sum_6_pcb
) |> select(-fi_id)


IM <- bind_rows(IMstructure , IM)

IM <- IM |> 
  filter(SamplingSchemeID %in% params$samplingscheme) 

write.csv(IM,
          file.path(folder, "IM.csv"),
          quote =FALSE,
          na = "",
          row.names = FALSE)

```

IM : `r nrow(IM)`

```{r}
#| label: Build_format
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'



files <- list.files(folder, pattern = "\\.csv$", full.names = TRUE)
order = unlist(mapply(grep, c("FI","SS","GM","GS", "IM"), MoreArgs = list(x=files)))
files <- files[order]
all_lines <- unlist(lapply(files, readLines))
con0 <- paste0("export_format/test_format_",
               paste(params$samplingscheme,collapse="_"),
               "0.csv")# Write to a new single file
# replace with new version if exists
num0 = 0
while (file.exists(con0)) {
  num = num0+1
  con0 <- paste0("export_format/test_format_",
               paste(params$samplingscheme,collapse="_"),num,
               ".csv")
}

# write the result as a single file in DATSU format

writeLines(all_lines, con=con0)

```

File exported : `r con0`